---
title: "Datacleaningcode_VA"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(tidyverse)
#here are functions
#for temperature

resultconvert_celsius<-function(dat, variable, unit, out_unit) {
  dat$ResultMeasureValue[which(dat$CharacteristicName == variable & dat$ResultMeasure.MeasureUnitCode == unit)] <- (dat$ResultMeasureValue[which(dat$CharacteristicName == variable & dat$ResultMeasure.MeasureUnitCode == unit)] -32)* 5/9
  
dat$ResultMeasure.MeasureUnitCode[which(dat$CharacteristicName == variable & dat$ResultMeasure.MeasureUnitCode == unit)] <- out_unit

return(dat)
  
}


detectconvert_celsius<-function(dat, variable, unit, out_unit) {
  dat$DetectionQuantitationLimitMeasure.MeasureValue[which(dat$CharacteristicName == variable & dat$DetectionQuantitationLimitMeasure.MeasureUnitCode == unit)] <- (dat$DetectionQuantitationLimitMeasure.MeasureValue[which(dat$CharacteristicName == variable & dat$DetectionQuantitationLimitMeasure.MeasureUnitCode == unit)] * -32)*5/9
  
dat$DetectionQuantitationLimitMeasure.MeasureUnitCode[which(dat$CharacteristicName == variable & dat$DetectionQuantitationLimitMeasure.MeasureUnitCode == unit)] <- out_unit

return(dat)
  
}

#for all other conversions

resultconvert<-function(dat, variable, unit, out_unit, conv_factor) {
  dat$ResultMeasureValue[which(dat$CharacteristicName == variable & dat$ResultMeasure.MeasureUnitCode == unit)] <- dat$ResultMeasureValue[which(dat$CharacteristicName == variable & dat$ResultMeasure.MeasureUnitCode == unit)] * conv_factor
  
dat$ResultMeasure.MeasureUnitCode[which(dat$CharacteristicName == variable & dat$ResultMeasure.MeasureUnitCode == unit)] <- out_unit

return(dat)
  
}


detectconvert<-function(dat, variable, unit, out_unit, conv_factor) {
  dat$DetectionQuantitationLimitMeasure.MeasureValue[which(dat$CharacteristicName == variable & dat$DetectionQuantitationLimitMeasure.MeasureUnitCode == unit)] <- dat$DetectionQuantitationLimitMeasure.MeasureValue[which(dat$CharacteristicName == variable & dat$DetectionQuantitationLimitMeasure.MeasureUnitCode == unit)] * conv_factor
  
dat$DetectionQuantitationLimitMeasure.MeasureUnitCode[which(dat$CharacteristicName == variable & dat$DetectionQuantitationLimitMeasure.MeasureUnitCode == unit)] <- out_unit

return(dat)
  
}



`%notin%` <- Negate(`%in%`)

VAdatPull<-read.csv("~/Lagos Limno/VA_alldata_1975_to_2020.csv")

names(VAdatPull)

summary(VAdatPull)

#removing data that have Y's instead of keeping data that have N's

read.csv("~/Lagos Limno/list of unique parameters to remove.csv") -> remove_params #update this


data<- VAdatPull[VAdatPull$CharacteristicName %notin% remove_params$Unique.parameters,]
data<-data %>% mutate(obs_id = seq(1, nrow(data), 1))


#tdata1<-data %>% filter(!is.na(ResultDepthHeightMeasure.MeasureValue) | !is.na(ActivityDepthHeightMeasure.MeasureValue)) #identified specified depth samples

#tdata2<-data %>% filter(is.na(ResultDepthHeightMeasure.MeasureValue)&is.na(ActivityDepthHeightMeasure.MeasureValue)) %>% filter(!is.na(ActivityTopDepthHeightMeasure.MeasureValue)) #these are potential integrated samples

#unique(tdata2$SampleCollectionMethod.MethodName)
#method_name<-c("Ambient Water Quality Monitoring QAPP and SOP")

#unique(tdata2$SampleCollectionMethod.MethodIdentifier)
#method_identifier<-c("")

#unique(tdata2$SampleCollectionMethod.MethodIdentifierContext)
#method_context<-c("")

#unique(tdata2$SampleCollectionEquipmentName)
#method_equipment<-c("")

#tdata2<-tdata2 %>% filter(SampleCollectionMethod.MethodName %in% method_name | 
                            #SampleCollectionMethod.MethodIdentifier %in% method_identifier|
                            #SampleCollectionMethod.MethodIdentifierContext %in% method_context|
                            #SampleCollectionEquipmentName %in% method_equipment)  #all grab samples

#tdata3<-data %>% filter(is.na(ResultDepthHeightMeasure.MeasureValue)) %>% filter(is.na(ActivityDepthHeightMeasure.MeasureValue)) #identifying grab samples

# unique(tdata3$SampleCollectionMethod.MethodName)
# method_name<-c("Ambient Water Quality Monitoring QAPP and SOP", "Grab", "Unspecified Standard Grab Sample Procedure", "Grab sample  (dip)", "Grab sample collection")
# 
# unique(tdata3$SampleCollectionMethod.MethodIdentifier)
# method_identifier<-c("000002", "NPS_GRAB", "Grab")
# 
# unique(tdata3$SampleCollectionMethod.MethodIdentifierContext)
# method_context<-c("USGS parameter code 82398")
# 
# unique(tdata3$SampleCollectionEquipmentName)
# method_equipment<-c("Bucket", "Open-Mouth Bottle", "Grab sample")
# 
# tdata3<-tdata3 %>% filter(SampleCollectionMethod.MethodName %in% method_name | 
#                             SampleCollectionMethod.MethodIdentifier %in% method_identifier|
#                             SampleCollectionMethod.MethodIdentifierContext %in% method_context|
#                             SampleCollectionEquipmentName %in% method_equipment)  #all grab samples
# 
# data2<-rbind(tdata1, tdata2, tdata3)
# rm(tdata1)
# rm(tdata2)
# rm(tdata3)

# tdata4<-data %>% filter(obs_id %notin% data2$obs_id)#created object to look at all data not in datata2
# 
# unique(tdata4$CharacteristicName)
# secchiparams<-c("Secchi, Horizontal Distance", "Depth, Secchi disk depth") #pulled out secchi data cuz it may not have depth data
# 
# tdata4<-tdata4 %>% filter(CharacteristicName %in% secchiparams) #filtered out only the secchi vars
# data<-rbind(data2, tdata4) #combined everything back together
# 
# #will need to check for each state (the depth data)
# 
# rm(tdata4)
# rm(data2)
# data<-data %>% select(-obs_id)
# rm(VAdatPull)
# unique(data$CharacteristicName)

str(data$ResultMeasureValue)
sum(is.na(data$ResultMeasureValue))
data1<-data %>% filter(!is.na(ResultMeasureValue))
data2<-data %>% filter(!is.na(DetectionQuantitationLimitMeasure.MeasureValue)) %>% filter(is.na(ResultMeasureValue))
data3<-rbind(data1, data2)
data<-data3

rm(data1, data2, data3)


#unique organizations
unique(data$OrganizationIdentifier) #19 organizations data
table(data$OrganizationFormalName) #tells us number of obvs in each organization




#looking at activity type codes
unique(data$ActivityTypeCode)
activitytype_keep<-c("Quality Control Sample-Field Replicate", "Quality Control Sample-Field Replicate", "Quality Control Sample-Field Blank", "Quality Control Sample-Equipment Blank", "Quality Control Sample-Equipment Blank", "Quality Control Field Sample Equipment Rinsate Blank", "Quality Control Sample-Reagent Blank", "Quality Control Field Replicate Msr/Obs")
data<-data[which(data$ActivityTypeCode %notin% activitytype_keep),]
unique(data$ActivityTypeCode)

#manually checking these columns to see if these is data we keep or can drop

unique(data$SubjectTaxonomicName) #expected it to be na - to make sure nothing we need is in there
unique(data$SampleTissueAnatomyName) #also NA's - expected

unique(data$ActivityMediaSubdivisionName)
#only keeping subdivision name with surface water and NA's - dropping everything else 
data<-data %>% filter(ActivityMediaSubdivisionName == "Surface Water" | is.na(ActivityMediaSubdivisionName))
unique(data$ActivityMediaSubdivisionName)

unique(data$ResultStatusIdentifier)

unique(data$ResultValueTypeName)
data<-data %>% filter(ResultValueTypeName == "Actual" | is.na(ResultValueTypeName))

unique(data$ResultDetectionConditionText)
#filtered out vars we don't need in resultsdetection condition text - did it in multistep because R doesn't handle multiple OR statements well


data<-data %>% filter(ResultDetectionConditionText != "Systematic Contamination" | is.na(ResultDetectionConditionText))
data<-data %>% filter(is.na(ResultDetectionConditionText) | ResultDetectionConditionText != "Not Reported")
unique(data$ResultDetectionConditionText)



unique(data$ResultSampleFractionText)

data<-data %>% filter(is.na(ResultSampleFractionText) | ResultSampleFractionText != "Bed Sediment")
unique(data$ResultSampleFractionText)



data<-data %>% select(-ActivityMediaName, -ActivityMediaSubdivisionName, -ActivityStartTime.Time, -ActivityStartTime.TimeZoneCode, -ActivityEndTime.Time, -ActivityEndTime.TimeZoneCode, -SampleAquifer, -HydrologicCondition, -HydrologicEvent, -SubjectTaxonomicName, -SampleTissueAnatomyName, -AnalysisStartDate, -PreparationStartDate)


#checking for depth conversions 
unique(data$ActivityDepthHeightMeasure.MeasureUnitCode) #has m, feet, ft, and NA - need to convert feet and ft to m 
unique(data$ActivityTopDepthHeightMeasure.MeasureUnitCode) #all NAs - no conversion needed
unique(data$ActivityBottomDepthHeightMeasure.MeasureUnitCode) #all NAs - no conversion needed
unique(data$ResultDepthHeightMeasure.MeasureUnitCode) #all NAs


temp<-filter(data, ActivityDepthHeightMeasure.MeasureUnitCode == "ft")
temp<-filter(data, ActivityDepthHeightMeasure.MeasureUnitCode == "feet")
temp<-filter(data, ActivityDepthHeightMeasure.MeasureUnitCode == "m")



##this method of conversion works
#temp<-data
#temp$ActivityDepthHeightMeasure.MeasureValue<-ifelse(temp$ActivityDepthHeightMeasure.MeasureUnitCode == "ft", temp$ActivityDepthHeightMeasure.MeasureValue*0.3048, temp$ActivityDepthHeightMeasure.MeasureValue)
#check to see if it worked  
#temp1<-filter(temp, ActivityDepthHeightMeasure.MeasureUnitCode == "ft")
#itworked


#use above method to convert values of "ft" and "feet" to "m"
#then set all values in ActivityDepthHeightMeasure.MeasureUnitCode to "m"

data$ActivityDepthHeightMeasure.MeasureValue<-ifelse(data$ActivityDepthHeightMeasure.MeasureUnitCode == "ft", data$ActivityDepthHeightMeasure.MeasureValue*0.3048, data$ActivityDepthHeightMeasure.MeasureValue)

data$ActivityDepthHeightMeasure.MeasureValue<-ifelse(data$ActivityDepthHeightMeasure.MeasureUnitCode == "feet", data$ActivityDepthHeightMeasure.MeasureValue*0.3048, data$ActivityDepthHeightMeasure.MeasureValue)


data$ActivityDepthHeightMeasure.MeasureUnitCode<-"m"

#check
unique(data$ActivityDepthHeightMeasure.MeasureUnitCode)

```

```{r}
#list of unique parameters
#unique(data$CharacteristicName)
#write.table(unique(data$CharacteristicName), "unique_parameters_VA.csv", row.names = FALSE)
#unique_parameters_VA<-read.csv("~/Lagos Limno/unique_parameters_VA.csv")
#data<-data[data$CharacteristicName %in% unique_parameters_VA$Unique_params,]
#unique(data$CharacteristicName)


```

```{r}

###creating new columns to preserve original pulled data

data$source_parameter <- data$CharacteristicName
data$source_value<- data$ResultMeasureValue
data$source_unit <- data$ResultMeasure.MeasureUnitCode
data$detectionlimit_legacy <- data$DetectionQuantitationLimitMeasure.MeasureValue
data$detectionlimit_unit_legacy <- data$DetectionQuantitationLimitMeasure.MeasureUnitCode

#spreadsheet for VA variables
#write.table(unique(data$CharacteristicName), "Va_vars.csv", row.names = FALSE)
```



```{r}
#secchi

unique(data$CharacteristicName)

Secchi_params<-c("Depth", "Depth, Secchi disk depth", "Secchi, Horizontal Distance") #values for secchi
data_secchi<-data[data$CharacteristicName%in%Secchi_params,]

data_secchi<-data[data$CharacteristicName%in%Secchi_params,] #want to match all records to the vector we created above which have TKN
unique(data_secchi$ResultSampleFractionText)
temp<-data_secchi %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep


###create dataset to check for non numeric objects in dataset
temp<-data_secchi %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_secchi<-data_secchi %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_secchi$ResultMeasure.MeasureUnitCode) #checked units
temp<-data_secchi[which(!is.na(data_secchi$DetectionQuantitationLimitMeasure.MeasureUnitCode) & data_secchi$ResultMeasure.MeasureUnitCode != data_secchi$DetectionQuantitationLimitMeasure.MeasureUnitCode),] #checks to see if detection limit units same as obvs units



###convert inches to m for obvservations
data_secchi$ResultMeasureValue[which(data_secchi$ResultMeasure.MeasureUnitCode == "in")] <- data_secchi$ResultMeasureValue[which(data_secchi$ResultMeasure.MeasureUnitCode == "in")] * 0.0254

data_secchi[(!is.na(data_secchi$DetectionQuantitationLimitMeasure.MeasureUnitCode) & data_secchi$ResultMeasure.MeasureUnitCode != data_secchi$DetectionQuantitationLimitMeasure.MeasureUnitCode),] #checks to see if detection limit units same as obvs units

#convert inches to m for detectionlimits
data_secchi$DetectionQuantitationLimitMeasure.MeasureValue[which(data_secchi$ResultMeasure.MeasureUnitCode == "in")] <- data_secchi$DetectionQuantitationLimitMeasure.MeasureValue[which(data_secchi$ResultMeasure.MeasureUnitCode == "in")] * 0.0254


####convert ft to m's for values

data_secchi$ResultMeasureValue[which(data_secchi$ResultMeasureValue == "ft")] <-
  data_secchi$ResultMeasureValue[which(data_secchi$ResultMeasureValue == "ft")] * 0.3048

#convert ft to m's for detection limits

data_secchi$DetectionQuantitationLimitMeasure.MeasureValue[which(data_secchi$DetectionQuantitationLimitMeasure.MeasureValue == "ft")]<-
  data_secchi$DetectionQuantitationLimitMeasure.MeasureValue[which(data_secchi$DetectionQuantitationLimitMeasure.MeasureValue == "ft")] * 0.3048

#then change all the units to m

data_secchi$ResultMeasure.MeasureUnitCode<-"m" #for values

data_secchi$DetectionQuantitationLimitMeasure.MeasureUnitCode<-"m"  #for detection limit
#check to see if it worked
unique(data_secchi$ResultMeasure.MeasureUnitCode)
unique(data_secchi$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_secchi$CharacteristicName<-"Secchi"  #set all secchi params as secchi

#done
```



```{r}
#calcium

Calcium_params<-c("Calcium") #only values for calcium
data_calcium<-data[data$CharacteristicName%in%Calcium_params,] #want to match all records to the vector we created above which have calcium
unique(data_calcium$ResultSampleFractionText)
temp<-data_calcium %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep



###create dataset to check for non numeric objects in dataset
temp<-data_calcium %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_calcium<-data_calcium %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_calcium$ResultMeasure.MeasureUnitCode) #checked units
unique(data_calcium$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#convert mg/l CaCO3 to mg/l Ca

data_calcium<-resultconvert(dat = data_calcium, variable = "Calcium", unit = "mg/l CaCO3", out_unit = "mg/l", conv_factor = 1/2.5)
data_calcium<-detectconvert(dat = data_calcium, variable = "Calcium", unit = "mg/l CacO3", out_unit = "mg/l", conv_factor = 1/2.5)

temp<- filter(data_calcium, ResultMeasure.MeasureUnitCode == "ueq/L") #check to see if ueq/l is for calcium or caco3 - it's for calcium

#need to convert ueq/l to meq/l (*1/000), then meq/l to mg/l (*20.04) - do this for resultconvert and detectconvert - 

data_calcium<-resultconvert(dat = data_calcium, variable = "Calcium", unit = "ueq/L", out_unit = "meq/l_converted", conv_factor = 1/1000)
data_calcium<-resultconvert(dat = data_calcium, variable = "Calcium", unit = "meq/l_converted", out_unit = "mg/l", conv_factor = 20.04)

#do the same for detect convert

data_calcium<-detectconvert(dat = data_calcium, variable = "Calcium", unit = "ueq/L", out_unit = "meq/l_converted", conv_factor = 1/1000)
data_calcium<-detectconvert(dat = data_calcium, variable = "Calcium", unit = "meq/l_converted", out_unit = "mg/l", conv_factor = 20.04)

#check to see everything converted 
temp<-data_calcium %>% filter(CharacteristicName == "Calcium")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_calcium$CharacteristicName<-"Calcium, dissolved"

#COMPLETED
```



```{r}
#aluminum, total

Aluminum_params<-c("Aluminum") #only values for aluminum
data_aluminum_total<-data[data$CharacteristicName%in%Aluminum_params,] #want to match all records to the vector we created above which have calcium
unique(data_aluminum_total$ResultSampleFractionText)
temp<-data_aluminum_total %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Total") #only want total aluminum
data_aluminum_total<-data_aluminum_total[which(data_aluminum_total$ResultSampleFractionText %in% fractionkeep | is.na(data_aluminum_total$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_aluminum_total$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_aluminum_total %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_aluminum_total<-data_aluminum_total %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_aluminum_total$ResultMeasure.MeasureUnitCode) #checked units
unique(data_aluminum_total$DetectionQuantitationLimitMeasure.MeasureUnitCode) #checked units - all NA's - won't have to convert any detection limits

#resultmeasurecode - all in ug/l which is lagos units - no conversion needed

data_aluminum_dissolved$CharacteristicName<-"Aluminum, total"


#completed


```


```{r}
#aluminum, dissolved

Aluminum_params<-c("Aluminum") #only values for aluminum
data_aluminum_dissolved<-data[data$CharacteristicName%in%Aluminum_params,] #want to match all records to the vector we created above which have calcium
unique(data_aluminum_dissolved$ResultSampleFractionText)
temp<-data_aluminum_dissolved %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved aluminum
data_aluminum_dissolved<-data_aluminum_dissolved[which(data_aluminum_dissolved$ResultSampleFractionText %in% fractionkeep | is.na(data_aluminum_dissolved$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_aluminum_dissolved$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_aluminum_dissolved %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_aluminum_dissolved<-data_aluminum_dissolved %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_aluminum_dissolved$ResultMeasure.MeasureUnitCode) #checked units
unique(data_aluminum_dissolved$DetectionQuantitationLimitMeasure.MeasureUnitCode) #checked units - all NA's - won't have to convert any detection limits

#resultmeasurecode - all in ug/l which is lagos units - no conversion needed

data_aluminum_dissolved$CharacteristicName<-"Aluminum, dissolved"

#completed
```




```{r}
#mercury_total


Mercury_params<-c("Mercury") #only values for mercury
data_mercury_total<-data[data$CharacteristicName%in%Mercury_params,] #want to match all records to the vector we created above which have calcium
unique(data_mercury_total$ResultSampleFractionText)
temp<-data_mercury_total %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Total") #only want total mercury
data_mercury_total<-data_mercury_total[which(data_mercury_total$ResultSampleFractionText %in% fractionkeep | is.na(data_mercury_total$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_mercury_total$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_mercury_total %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #found 7 objects - all NA's - safe to convert
data_mercury_total<-data_mercury_total %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_mercury_total$ResultMeasure.MeasureUnitCode) #checked units
unique(data_mercury_total$DetectionQuantitationLimitMeasure.MeasureUnitCode)#checked units - same as resultmeasurevalues

#need to convert ng/l to ug/l

data_mercury_total<-resultconvert(dat = data_mercury_total, variable = "Mercury", unit = "ug/l", out_unit = "ng/l", conv_factor = 0.001)
data_mercury_total<-detectconvert(dat = data_mercury_total, variable = "Mercury", unit = "ug/l", out_unit = "ng/l", conv_factor = 0.001)

#check to make sure they worked
unique(data_mercury_total$ResultMeasure.MeasureUnitCode) #checked units
unique(data_mercury_total$DetectionQuantitationLimitMeasure.MeasureUnitCode)#checked units - same as resultmeasurevalues

data_mercury_total$CharacteristicName<-"Mercury, total"


#completed

```

```{r}
#mercury_dissolved


Mercury_params<-c("Mercury") #only values for mercury
data_mercury_dissolved<-data[data$CharacteristicName%in%Mercury_params,] #want to match all records to the vector we created above which have calcium
unique(data_mercury_dissolved$ResultSampleFractionText)
temp<-data_mercury_dissolved %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved mercury
data_mercury_dissolved<-data_mercury_dissolved[which(data_mercury_dissolved$ResultSampleFractionText %in% fractionkeep | is.na(data_mercury_dissolved$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_mercury_dissolved$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_mercury_dissolved %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #found 7 objects - all NA's - safe to convert
data_mercury_dissolved<-data_mercury_dissolved %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_mercury_dissolved$ResultMeasure.MeasureUnitCode) #checked units
unique(data_mercury_dissolved$DetectionQuantitationLimitMeasure.MeasureUnitCode)#checked units - same as resultmeasurevalues

#already in ng/l - no conversion needed

data_mercury_dissolved$CharacteristicName<-"Mercury, total dissolved"
#completed
```


```{r}
#Magnesium

Magnesium_params<-c("Magnesium") #only values for magnesium
data_magnesium<-data[data$CharacteristicName%in%Magnesium_params,] #want to match all records to the vector we created above which have calcium
unique(data_magnesium$ResultSampleFractionText)
temp<-data_magnesium %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved mercury
data_magnesium<-data_magnesium[which(data_magnesium$ResultSampleFractionText %in% fractionkeep | is.na(data_magnesium$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_magnesium$ResultSampleFractionText)



###create dataset to check for non numeric objects in dataset
temp<-data_magnesium %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_magnesium<-data_magnesium %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_magnesium$ResultMeasure.MeasureUnitCode) #checked units
unique(data_magnesium$DetectionQuantitationLimitMeasure.MeasureUnitCode) #checked units

#need to convert ueq/l to mg/l - 2 step process - need to do this for results and detectionlimits


data_magnesium<-resultconvert(dat = data_magnesium, variable = "Magnesium", unit = "ueq/L", out_unit = "meq/l_converted", conv_factor = 1/1000)
data_magnesium<-resultconvert(dat = data_magnesium, variable = "Magnesium", unit = "meq/l_converted", out_unit = "mg/l", conv_factor = 20.04)

#do the same for detect convert

data_magnesium<-detectconvert(dat = data_magnesium, variable = "Magnesium", unit = "ueq/L", out_unit = "meq/l_converted", conv_factor = 1/1000)
data_magnesium<-detectconvert(dat = data_magnesium, variable = "Magnesium", unit = "meq/l_converted", out_unit = "mg/l", conv_factor = 20.04)

#check to see everything converted 
temp<-data_magnesium %>% filter(CharacteristicName == "Magnesium")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_magnesium$CharacteristicName<-"Magnesium, dissolved"

#completed
```

```{r}
#selenium_total

Selenium_params<-c("Selenium") #only values for selenium
data_selenium_total<-data[data$CharacteristicName%in%Selenium_params,] #want to match all records to the vector we created above which have calcium
unique(data_selenium_total$ResultSampleFractionText)
temp<-data_selenium_total %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Total") #only want total
data_selenium_total<-data_selenium_total[which(data_selenium_total$ResultSampleFractionText %in% fractionkeep | is.na(data_selenium_total$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_selenium_total$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_selenium_total %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change # 21 NA - can change
data_selenium_total<-data_selenium_total %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_selenium_total$ResultMeasure.MeasureUnitCode)#checked units
unique(data_selenium_total$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#units matched - all ug/l - this is lagos units - no conversions needed

data_selenium_total$CharacteristicName<-"Selenium, total"

#completed
```

```{r}
#selenium_dissolved

Selenium_params<-c("Selenium") #only values for selenium
data_selenium_dissolved<-data[data$CharacteristicName%in%Selenium_params,] #want to match all records to the vector we created above which have calcium
unique(data_selenium_dissolved$ResultSampleFractionText)
temp<-data_selenium_dissolved %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved
data_selenium_dissolved<-data_selenium_dissolved[which(data_selenium_dissolved$ResultSampleFractionText %in% fractionkeep | is.na(data_selenium_dissolved$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_selenium_dissolved$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_selenium_dissolved %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change 
data_selenium_dissolved<-data_selenium_dissolved %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_selenium_dissolved$ResultMeasure.MeasureUnitCode)#checked units
unique(data_selenium_dissolved$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#units matched - all ug/l - this is lagos units - no conversions needed

data_selenium_dissolved$CharacteristicName<-"Selenium, dissolved"


```



```{r}
#arsenic_total

Arsenic_params<-c("Arsenic") #only values for arsenic
data_arsenic_total<-data[data$CharacteristicName%in%Arsenic_params,] #want to match all records to the vector we created above which have calcium
unique(data_arsenic_total$ResultSampleFractionText)
temp<-data_arsenic_total %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Total") #only want total
data_arsenic_total<-data_arsenic_total[which(data_arsenic_total$ResultSampleFractionText %in% fractionkeep | is.na(data_arsenic_total$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_arsenic_total$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_arsenic_total %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #7 NA's - can change
data_arsenic_total<-data_arsenic_total %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_arsenic_total$ResultMeasure.MeasureUnitCode)#checked units
unique(data_arsenic_total$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#already in ug/l - lagos units 

data_arsenic_total$CharacteristicName<-"Arsenic, total"

#completed

```


```{r}
#arsenic_dissolved

Arsenic_params<-c("Arsenic") #only values for arsenic
data_arsenic_dissolved<-data[data$CharacteristicName%in%Arsenic_params,] #want to match all records to the vector we created above which have calcium
unique(data_arsenic_dissolved$ResultSampleFractionText)
temp<-data_arsenic_dissolved %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved
data_arsenic_dissolved<-data_arsenic_dissolved[which(data_arsenic_dissolved$ResultSampleFractionText %in% fractionkeep | is.na(data_arsenic_dissolved$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_arsenic_dissolved$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_arsenic_dissolved %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #7 NA's - can change
data_arsenic_dissolved<-data_arsenic_dissolved %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_arsenic_dissolved$ResultMeasure.MeasureUnitCode)#checked units
unique(data_arsenic_dissolved$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#already in ug/l - lagos units 

data_arsenic_dissolved$CharacteristicName<-"Arsenic, dissolved"

#completed

```


```{r}
#atrazine_total #no observations for total

Atrazine_params<-c("Atrazine") #only values for atrazine
data_atrazine_total<-data[data$CharacteristicName%in%Atrazine_params,] #want to match all records to the vector we created above which have calcium
unique(data_atrazine_total$ResultSampleFractionText)
temp<-data_atrazine_total %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Total") #only want total
data_atrazine_total<-data_atrazine_total[which(data_atrazine_total$ResultSampleFractionText %in% fractionkeep | is.na(data_atrazine_total$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_atrazine_total$ResultSampleFractionText)



###create dataset to check for non numeric objects in dataset
temp<-data_atrazine_total %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #3 NA's - can change
data_atrazine_total<-data_atrazine_total %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_atrazine_total$ResultMeasure.MeasureUnitCode)#checked units
unique(data_atrazine_total$DetectionQuantitationLimitMeasure.MeasureUnitCode)
#both in ug/l - lagos units - no conversion needed


#completed
```

```{r}
#atrazine_dissolved 

Atrazine_params<-c("Atrazine") #only values for atrazine
data_atrazine_dissolved<-data[data$CharacteristicName%in%Atrazine_params,] #want to match all records to the vector we created above which have calcium
unique(data_atrazine_dissolved$ResultSampleFractionText)
temp<-data_atrazine_dissolved %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved
data_atrazine_dissolved<-data_atrazine_dissolved[which(data_atrazine_dissolved$ResultSampleFractionText %in% fractionkeep | is.na(data_atrazine_dissolved$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_atrazine_dissolved$ResultSampleFractionText)



###create dataset to check for non numeric objects in dataset
temp<-data_atrazine_dissolved %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #3 NA's - can change
data_atrazine_dissolved<-data_atrazine_dissolved %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_atrazine_dissolved$ResultMeasure.MeasureUnitCode)#checked units
unique(data_atrazine_dissolved$DetectionQuantitationLimitMeasure.MeasureUnitCode)
#both in ug/l - lagos units - no conversion needed

data_atrazine_dissolved$CharacteristicName<-"Atrazine, dissolved"

```


```{r}
#ecoli
unique(data$CharacteristicName)

Ecoli_params<-c("Escherichia coli") #only values for Escherichia coli
data_ecoli<-data[data$CharacteristicName%in%Ecoli_params,] #want to match all records to the vector we created above which have calcium
unique(data_ecoli$ResultSampleFractionText)
temp<-data_ecoli %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep



###create dataset to check for non numeric objects in dataset
temp<-data_ecoli %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #1 NA - can change
data_ecoli<-data_ecoli %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_ecoli$ResultMeasure.MeasureUnitCode) #checked units
unique(data_ecoli$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#already in lagos units - MNP/100ml and cfu/100ml 

data_ecoli$CharacteristicName<-"E.coli"

#completed
```

```{r}
#silica

Silica_params<-c("Silica") #only values for silica
data_silica<-data[data$CharacteristicName%in%Silica_params,] #want to match all records to the vector we created above which have calcium
unique(data_silica$ResultSampleFractionText)
temp<-data_silica %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved
data_silica<-data_silica[which(data_silica$ResultSampleFractionText %in% fractionkeep | is.na(data_silica$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_silica$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_silica %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_silica<-data_silica %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_silica$ResultMeasure.MeasureUnitCode) #checked units
unique(data_silica$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#lagos units for silica is mg/l - need to convert units for resultmeasurevalue from umol/L to mg/l = *0.1151


#data_silica<-resultconvert(dat = data_silica, variable = "Silica", unit = "umol/L", out_unit = "mg/l", conv_factor = 0.1151)
#data_silica<-detectconvert(dat = data_silica, variable = "Silica", unit = "umol/L", out_unit = "mg/l", conv_factor = 0.1151)

#check to see everything converted 
#temp<-data_silica %>% filter(CharacteristicName == "Silica")
#unique(temp$ResultMeasure.MeasureUnitCode)
#unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_silica$CharacteristicName<-"Silica, dissolved"

#completed

```

```{r}
#chloride

Chloride_params<-c("Chloride") #only values for chloride
data_chloride<-data[data$CharacteristicName%in%Chloride_params,] #want to match all records to the vector we created above which have calcium
unique(data_chloride$ResultSampleFractionText)
temp<-data_chloride %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved
data_chloride<-data_chloride[which(data_chloride$ResultSampleFractionText %in% fractionkeep | is.na(data_chloride$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_chloride$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_chloride %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_chloride<-data_chloride %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_chloride$ResultMeasure.MeasureUnitCode) #checked units
unique(data_chloride$DetectionQuantitationLimitMeasure.MeasureUnitCode)



#need to convert ueq/l to meq/l (*1/000), then meq/l to mg/l (*20.04) - do this for resultconvert and detectconvert - 

data_chloride<-resultconvert(dat = data_chloride, variable = "Chloride", unit = "ueq/L", out_unit = "meq/l_converted", conv_factor = 1/1000)
data_chloride<-resultconvert(dat = data_chloride, variable = "Chloride", unit = "meq/l_converted", out_unit = "mg/l", conv_factor = 20.04)

#do the same for detect convert

data_chloride<-detectconvert(dat = data_chloride, variable = "Chloride", unit = "ueq/L", out_unit = "meq/l_converted", conv_factor = 1/1000)
data_chloride<-detectconvert(dat = data_chloride, variable = "Chloride", unit = "meq/l_converted", out_unit = "mg/l", conv_factor = 20.04)

#check to see everything converted 
temp<-data_chloride %>% filter(CharacteristicName == "Chloride")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_chloride$CharacteristicName<-"Chloride, dissolved"

```



```{r}
#oxygen

Oxygen_params<-c("Oxygen", "Dissolved oxygen (DO)") #only values for oxygen
data_oxygen<-data[data$CharacteristicName%in%Oxygen_params,] #want to match all records to the vector we created above which have oxygen
unique(data_oxygen$ResultSampleFractionText)
temp<-data_oxygen %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
#both types are dissolved oxygen - convert all characteristic names to oxygen
data_oxygen$CharacteristicName<-"Oxygen"
unique(data_oxygen$CharacteristicName)



###create dataset to check for non numeric objects in dataset
temp<-data_oxygen %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #1 NA observation - can change
data_oxygen<-data_oxygen %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_oxygen$ResultMeasure.MeasureUnitCode) #checked units
unique(data_oxygen$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#resultmeasurevalue has ppm, mg/l, and % saturation - need all in mg/l - detection limits all in mg/l - no need to change 

temp<-filter(data_oxygen, ResultMeasure.MeasureUnitCode == "% saturatn")
#1ppm = 1mg/l, and 1 DO% saturation = 0.08 mg/l

data_oxygen<-resultconvert(dat = data_oxygen, variable = "Oxygen", unit = "ppm", out_unit = "mg/l", conv_factor = 1)
data_oxygen<-resultconvert(dat = data_oxygen, variable = "Oxygen", unit = "% saturatn", out_unit = "mg/l", conv_factor = 0.08)
data_oxygen<-resultconvert(dat = data_oxygen, variable = "Oxygen", unit = "%", out_unit = "mg/l", conv_factor = 0.08)

#check to see everything converted 
temp<-data_oxygen %>% filter(CharacteristicName == "Oxygen")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_oxygen$CharacteristicName<-"Oxygen, dissolved"

#completed

```


```{r}
#chlorophyll

unique(data$CharacteristicName)


Chla_params<-c("Chlorophyll a, corrected for pheophytin", "Chlorophyll a", "Chlorophyll a, uncorrected for pheophytin", "Chlorophyll") #only values for Chl
data_chl<-data[data$CharacteristicName%in%Chla_params,] #want to match all records to the vector we created above which have chl
unique(data_chl$ResultSampleFractionText)
temp<-data_chl %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep



###create dataset to check for non numeric objects in dataset
temp<-data_chl %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #2 NA's - can change
data_chla<-data_chl%>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_chl$ResultMeasure.MeasureUnitCode) #checked units
unique(data_chl$DetectionQuantitationLimitMeasure.MeasureUnitCode)
#results and detection limit both in ug/l - Lagos units - no conversions needed

data_chl$CharacteristicName<-"Chlorophyll a"

#completed

```





```{r}
# pH

pH_params<-c("pH") #only values for pH
data_pH<-data[data$CharacteristicName%in%pH_params,] #want to match all records to the vector we created above which have pH
unique(data_pH$ResultSampleFractionText)
temp<-data_pH %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep



###create dataset to check for non numeric objects in dataset
temp<-data_pH %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_pH<-data_pH %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_pH$ResultMeasure.MeasureUnitCode) #checked units
unique(data_pH$DetectionQuantitationLimitMeasure.MeasureUnitCode)

temp<-filter(data_pH, DetectionQuantitationLimitMeasure.MeasureUnitCode == "None")

#convert text - from std units to pH and none to pH - (for detection limits) data is correct, but mislabelled - same for resultsmeasurevalue

data_pH$DetectionQuantitationLimitMeasure.MeasureUnitCode<-"pH"
unique(data_pH$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_pH$ResultMeasure.MeasureUnitCode<-"pH"
unique(data_pH$ResultMeasure.MeasureUnitCode)


data_pH$CharacteristicName<-"pH, equilibrated"

#completed
```

```{r}
#temperature

Temp_params<-c("Temperature, water") #only values for water temp
data_temp<-data[data$CharacteristicName%in%Temp_params,] #want to match all records to the vector we created above which have calcium
unique(data_temp$ResultSampleFractionText)
temp<-data_temp %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep



###create dataset to check for non numeric objects in dataset
temp<-data_temp %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_temp<-data_temp %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_temp$ResultMeasure.MeasureUnitCode) #checked units
unique(data_temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#temp is in deg C - already lagos units - no need to change

data_temp$CharacteristicName<-"Temperature"

#completed
```



```{r}
#sulfate
Sulfate_params<-c("Sulfate") #only values for sulfate
data_sulfate<-data[data$CharacteristicName%in%Sulfate_params,] #want to match all records to the vector we created above which have calcium
unique(data_sulfate$ResultSampleFractionText)
temp<-data_sulfate %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved
data_sulfate<-data_sulfate[which(data_sulfate$ResultSampleFractionText %in% fractionkeep | is.na(data_sulfate$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_sulfate$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_sulfate %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_sulfate<-data_sulfate %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_sulfate$ResultMeasure.MeasureUnitCode) #checked units
unique(data_sulfate$DetectionQuantitationLimitMeasure.MeasureUnitCode)



#need to convert ueq/l to meq/l (*1/000), then meq/l to mg/l (*20.04) - do this for resultconvert and detectconvert - 

data_sulfate<-resultconvert(dat = data_sulfate, variable = "Sulfate", unit = "ueq/L", out_unit = "meq/l_converted", conv_factor = 1/1000)
data_sulfate<-resultconvert(dat = data_sulfate, variable = "Sulfate", unit = "meq/l_converted", out_unit = "mg/l", conv_factor = 20.04)

#do the same for detect convert

data_sulfate<-detectconvert(dat = data_sulfate, variable = "Sulfate", unit = "ueq/L", out_unit = "meq/l_converted", conv_factor = 1/1000)
data_sulfate<-detectconvert(dat = data_sulfate, variable = "Sulfate", unit = "meq/l_converted", out_unit = "mg/l", conv_factor = 20.04)

#check to see everything converted 
temp<-data_sulfate %>% filter(CharacteristicName == "Sulfate")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_sulfate$CharacteristicName<-"Sulfate, dissolved"


```

```{r}
#turbidity

Turbidity_params<-c("Turbidity") #only values for turbidity
data_turb<-data[data$CharacteristicName%in%Turbidity_params,] #want to match all records to the vector we created above which have calcium
unique(data_turb$ResultSampleFractionText)
temp<-data_turb %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep



###create dataset to check for non numeric objects in dataset
temp<-data_turb %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_turb<-data_turb %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_turb$ResultMeasure.MeasureUnitCode) #checked units
unique(data_turb$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#detection limit is in NTU - lagos units - don't need to change
#resultmeasurevalue has FTU and FNU - need to convert to NTU

#FTU= NTU, and FNU= NTU ( suggested at similar/comparable - no direct conversion)
#converting both readings to NTU

data_turb$ResultMeasure.MeasureUnitCode<-"NTU"
unique(data_turb$ResultMeasure.MeasureUnitCode)

data_turb$CharacteristicName<-"Turbidity"

#completed

```

```{r}
#conductivity

Cond_params<-c("Specific conductance") #only values for conductivity
data_cond<-data[data$CharacteristicName%in%Cond_params,] #want to match all records to the vector we created above which have calcium
unique(data_cond$ResultSampleFractionText)
temp<-data_cond %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep



###create dataset to check for non numeric objects in dataset
temp<-data_cond %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_cond<-data_cond %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_cond$ResultMeasure.MeasureUnitCode) #checked units
unique(data_cond$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#detectionlimit already in lagos units - uS/cm - need to convert resultmeasure from umho/cm to us/cm - conversion = 1:1 - will change text(unit code)- no conversion needed

data_cond<-resultconvert(dat = data_cond, variable = "Specific conductance", unit = "umho/cm", out_unit = "uS/cm", conv_factor = 1)
unique(data_cond$ResultMeasure.MeasureUnitCode)

#changing us/cm @25C to us/cm for resultmeasure and detection limits

data_cond<-resultconvert(dat = data_cond, variable = "Specific conductance", unit = "uS/cm @25C", out_unit = "uS/cm", conv_factor = 1)
data_cond<-detectconvert(dat = data_cond, variable = "Specific conductance", unit = "uS/cm @25C", out_unit = "uS/cm", conv_factor = 1)


#check to see if properly converted

unique(data_cond$ResultMeasure.MeasureUnitCode)
unique(data_cond$DetectionQuantitationLimitMeasure.MeasureUnitCode)


data_cond$CharacteristicName<-"Conductivity"

#completed
```


```{r}
#Total carbon (TC)

Carbon_params<-c("Carbon", "Organic carbon") #only values for carbon
data_TC<-data[data$CharacteristicName%in%Carbon_params,] #want to match all records to the vector we created above which have calcium
unique(data_TC$ResultSampleFractionText)
temp<-data_TC %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Total") #only want dissolved organic carbon - not total or suspended
data_TC<-data_TC[which(data_TC$ResultSampleFractionText %in% fractionkeep | is.na(data_TC$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_TC$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_TC %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_TC<-data_TC %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_TC$ResultMeasure.MeasureUnitCode) #checked units
unique(data_TC$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#units are mg/l - this is lagos units - no need for conversion

data_TC$CharacteristicName<-"Carbon, total organic"

#completed
```


```{r}
#Total dissolved carbon (TDC)

Carbon_params<-c("Carbon", "Organic carbon") #only values for carbon
data_TDC<-data[data$CharacteristicName%in%Carbon_params,] #want to match all records to the vector we created above which have calcium
unique(data_TDC$ResultSampleFractionText)
temp<-data_TDC %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved") #only want dissolved organic carbon - not total or suspended
data_TDC<-data_TDC[which(data_TDC$ResultSampleFractionText %in% fractionkeep | is.na(data_TDC$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_TDC$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_TDC %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_TDC<-data_TDC %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_TDC$ResultMeasure.MeasureUnitCode) #checked units
unique(data_TDC$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#units are mg/l - this is lagos units - no need for conversion

data_TDC$CharacteristicName<-"Carbon, dissolved organic"

#completed
```



```{r}
#alkalinity


Alk_params<-c("Alkalinity", "Alkalinity, total") #only values for alkalinity
data_alk<-data[data$CharacteristicName%in%Alk_params,] #want to match all records to the vector we created above which have calcium
unique(data_alk$ResultSampleFractionText) #keeping both total and dissolved
temp<-data_alk %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep



###create dataset to check for non numeric objects in dataset
temp<-data_alk %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change
data_alk<-data_alk %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_alk$ResultMeasure.MeasureUnitCode) #checked units
unique(data_alk$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#need to convert mg/l CaCO3 to ueq/l - multiply caco3*19.98, will be converting mg/l to ueql the same way - *19.98


temp<-filter(data_alk, ResultMeasure.MeasureUnitCode == "mg/l") 
unique(temp$CharacteristicName) #all alkalinity totals have mg/l - will be converting alk total and alk seperately

#converting alk total from mg/l to ueq/l
data_alk<-resultconvert(dat = data_alk, variable = "Alkalinity, total", unit = "mg/l", out_unit = "ueq/l", conv_factor = 19.98)
unique(data_alk$ResultMeasure.MeasureUnitCode)

temp<-filter(data_alk, ResultMeasure.MeasureUnitCode == "mg/l CaCO3") #these are also alk total - so will be converting alk total again

#converting alk total from mg/l caco3 to ueq/l
data_alk<-resultconvert(dat = data_alk, variable = "Alkalinity, total", unit = "mg/l CaCO3", out_unit = "ueq/l", conv_factor = 19.98)
unique(data_alk$ResultMeasure.MeasureUnitCode)

#convert mg/l caco3 to ueq/l for detection limits

data_alk<-detectconvert(dat = data_alk, variable = "Alkalinity", unit = "mg/l CaCO3", out_unit = "ueq/L", conv_factor = 19.98)
unique(data_alk$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_alk$CharacteristicName<-"Alkalinity"


#completed

```

```{r}
#TSS

Tss_params<-c("Total suspended solids", "Total solids", "Total dissolved solids") #only values for tss
data_tss<-data[data$CharacteristicName%in%Tss_params,] #want to match all records to the vector we created above which have calcium
unique(data_tss$ResultSampleFractionText)
temp<-data_tss %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
#we only want suspended and total
fractionkeep<-c("Suspended", "Total") 
data_tss<-data_tss[which(data_tss$ResultSampleFractionText %in% fractionkeep | is.na(data_tss$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_tss$ResultSampleFractionText)




###create dataset to check for non numeric objects in dataset
temp<-data_tss %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #3 NA's - can change
data_tss<-data_tss %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_tss$ResultMeasure.MeasureUnitCode) #checked units
unique(data_tss$DetectionQuantitationLimitMeasure.MeasureUnitCode)

temp<-filter(data_tss, ResultMeasure.MeasureUnitCode == "%") #2 weird values measured in perentage rather than mg/l - drop these 2 from records
data_tss<-filter(data_tss, ResultMeasure.MeasureUnitCode != "%")
temp<-filter(data_tss, ResultMeasure.MeasureUnitCode == "%") #checked and it worked  

unique(data_tss$ResultMeasure.MeasureUnitCode) #checked units
unique(data_tss$DetectionQuantitationLimitMeasure.MeasureUnitCode)  
  
#units in mg/l - lagos units - no conversions needed

data_tss$CharacteristicName<-"Solids, total suspended"

#completed

```


```{r}
###DIP (dissolved inorganic phosphorus)

unique(data$CharacteristicName)

DIP_params<-c("Orthophosphate") #only values for DIP
data_DIP<-data[data$CharacteristicName%in%DIP_params,] #want to match all records to the vector we created above which have DIP
unique(data_DIP$ResultSampleFractionText)
temp<-data_DIP %>% filter(is.na(ResultSampleFractionText))
unique(temp$CharacteristicName) #this characteristic name is typically associated with dissolved 
fractionkeep<-c("Dissolved")
data_DIP<-data_DIP[which(data_DIP$ResultSampleFractionText %in% fractionkeep | is.na(data_DIP$ResultSampleFractionText)),]
unique(data_DIP$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_DIP %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values - if NA's not losing any other data - can change
#1905 NA's - can change
data_DIP<-data_DIP %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA


unique(data_DIP$ResultMeasure.MeasureUnitCode) #check units
unique(data_DIP$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#convert orthophosphate mg/l to ug/L #assuming it is PO4 even though it doesnt mention - we noticed it's dissolved - we;re assumming

data_DIP<-resultconvert(dat = data_DIP, variable = "Orthophosphate", unit = "mg/l", out_unit = "ug/L", conv_factor = 326.1)


#convert orthophosphate mg/l as P to ug/L

data_DIP<-resultconvert(dat = data_DIP, variable = "Orthophosphate", unit = "mg/l as P", out_unit = "ug/L", conv_factor = 1000)
data_DIP<-detectconvert(dat = data_DIP, variable = "Orthophosphate", unit = "mg/l as P", out_unit = "ug/L", conv_factor = 1000)


#convert orthophosphate mg/l as PO4 to ug/L
data_DIP<-resultconvert(dat = data_DIP, variable = "Orthophosphate", unit = "mg/l asPO4", out_unit = "ug/L", conv_factor = 0.3261*1000)
data_DIP<-detectconvert(dat = data_DIP, variable = "Orthophosphate", unit = "mg/l asPO4", out_unit = "ug/L", conv_factor = 0.3261*1000)


temp<-filter(data_DIP, ResultMeasure.MeasureUnitCode == "ueq/L")
temp<-filter(data_DIP, DetectionQuantitationLimitMeasure.MeasureUnitCode == "ueq/L")
#only 21 observations of each - drop them

data_DIP<-filter(data_DIP, ResultMeasure.MeasureUnitCode != "ueq/L")

unique(data_DIP$ResultMeasure.MeasureUnitCode) #check units
unique(data_DIP$DetectionQuantitationLimitMeasure.MeasureUnitCode)


data_DIP$CharacteristicName<-"Phosphorus, soluable reactive orthophosphate"

#COMPLETED

```


```{r}

#total phosphorus

unique(data$CharacteristicName)

TP_params<-c("Phosphorus") #only values for total phosphorus
data_TP<-data[data$CharacteristicName%in%TP_params,] #want to match all records to the vector we created above which have TP
unique(data_TP$ResultSampleFractionText)
temp<-data_TP %>% filter(is.na(ResultSampleFractionText))
fractionkeep<-c("Total")
data_TP<-data_TP[which(data_TP$ResultSampleFractionText %in% fractionkeep | is.na(data_TP$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_TP$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_TP %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values - if NA's not losing any other data - can change
data_TP<-data_TP %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_TP$ResultMeasure.MeasureUnitCode) #check units
unique(data_TP$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#convert mg/l PO4, mg/l as P, and mg/l to ug/l for resultsvalue and detectionlimits

data_TP<-resultconvert(dat = data_TP, variable = "Phosphorus", unit = "mg/l as P", out_unit = "ug/L", conv_factor = 1000)
data_TP<-detectconvert(dat = data_TP, variable = "Phosphorus", unit = "mg/l as P", out_unit = "ug/L", conv_factor = 1000)

data_TP<-resultconvert(dat = data_TP, variable = "Phosphorus", unit = "mg/l", out_unit = "ug/L", conv_factor = 1000)
data_TP<-detectconvert(dat = data_TP, variable = "Phosphorus", unit = "mg/l", out_unit = "ug/L", conv_factor = 1000)

data_TP<-resultconvert(dat = data_TP, variable = "Phosphorus", unit = "mg/l PO4", out_unit = "ug/L", conv_factor = 0.3261*1000)


unique(data_TP$ResultMeasure.MeasureUnitCode) #check units
unique(data_TP$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_TP$CharacteristicName<-"Phosphorus, total"


#complete

```


```{r}
#total dissolved phosphorus

unique(data$CharacteristicName)

TDP_params<-c("Phosphorus") #only values for total phosphorus
data_TDP<-data[data$CharacteristicName%in%TDP_params,] #want to match all records to the vector we created above which have TP
unique(data_TDP$ResultSampleFractionText)
temp<-data_TDP %>% filter(is.na(ResultSampleFractionText))
fractionkeep<-c("Dissolved")
data_TDP<-data_TDP[which(data_TDP$ResultSampleFractionText %in% fractionkeep | is.na(data_TDP$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_TDP$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_TDP %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values - if NA's not losing any other data - can change
data_TDP<-data_TDP %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_TDP$ResultMeasure.MeasureUnitCode) #check units
unique(data_TDP$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#convert mg/l as P and mg/l to ug/l for resultsvalue and detectionlimits

data_TDP<-resultconvert(dat = data_TDP, variable = "Phosphorus", unit = "mg/l as P", out_unit = "ug/L", conv_factor = 1000)
data_TDP<-detectconvert(dat = data_TDP, variable = "Phosphorus", unit = "mg/l as P", out_unit = "ug/L", conv_factor = 1000)

data_TDP<-resultconvert(dat = data_TDP, variable = "Phosphorus", unit = "mg/l", out_unit = "ug/L", conv_factor = 1000)
data_TDP<-detectconvert(dat = data_TDP, variable = "Phosphorus", unit = "mg/l", out_unit = "ug/L", conv_factor = 1000)


unique(data_TDP$ResultMeasure.MeasureUnitCode) #check units
unique(data_TDP$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_TDP$CharacteristicName<-"Phosphorus, total dissolved"


#complete

```


```{r}
#TKN #kjeldahl 

unique(data$CharacteristicName)

TKN_params<-c("Kjeldahl nitrogen") #only values for TKN
data_TKN<-data[data$CharacteristicName%in%TKN_params,] #want to match all records to the vector we created above which have TKN
unique(data_TKN$ResultSampleFractionText)
temp<-data_TKN %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Total")
data_TKN<-data_TKN[which(data_TKN$ResultSampleFractionText %in% fractionkeep | is.na(data_TKN$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_TKN$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_TKN %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change - 52 NA's - can change
data_TKN<-data_TKN %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_TKN$ResultMeasure.MeasureUnitCode)
unique(data_TKN$DetectionQuantitationLimitMeasure.MeasureUnitCode)#checked units



#WORKING WITH Kjeldahl nitrogen

temp<-data_TKN %>% filter(CharacteristicName == "Kjeldahl nitrogen") %>% filter(is.na(ResultMeasure.MeasureUnitCode)) %>% filter(!is.na(ResultMeasureValue))


temp<-data_TKN %>% filter(CharacteristicName == "Kjeldahl nitrogen")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#We have mg/l and mg/l as N - Need to make conversion for both into ug/L as N

data_TKN<-resultconvert(dat = data_TKN, variable = "Kjeldahl nitrogen", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)
data_TKN<-detectconvert(dat = data_TKN, variable = "Kjeldahl nitrogen", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)

data_TKN<-resultconvert(dat = data_TKN, variable = "Kjeldahl nitrogen", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)
data_TKN<-detectconvert(dat = data_TKN, variable = "Kjeldahl nitrogen", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)

#check to see everything converted 
temp<-data_TKN %>% filter(CharacteristicName == "Kjeldahl nitrogen")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_TKN$CharacteristicName<-"Nitrogen, total Kjeldahl"

#COMPLETED


```


```{r}
#TDKN #kjeldahl 

unique(data$CharacteristicName)

tdkn_params<-c("Kjeldahl nitrogen") #only values for tdkn
data_tdkn<-data[data$CharacteristicName%in%tdkn_params,] #want to match all records to the vector we created above which have tdkn
unique(data_tdkn$ResultSampleFractionText)
temp<-data_tdkn %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved")
data_tdkn<-data_tdkn[which(data_tdkn$ResultSampleFractionText %in% fractionkeep | is.na(data_tdkn$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_tdkn$ResultSampleFractionText)


###create dataset to check for non numeric objects in dataset
temp<-data_tdkn %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change - 52 NA's - can change
data_tdkn<-data_tdkn %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_tdkn$ResultMeasure.MeasureUnitCode)
unique(data_tdkn$DetectionQuantitationLimitMeasure.MeasureUnitCode)#checked units



#WORKING WITH Kjeldahl nitrogen


#We have mg/l and mg/l as N - Need to make conversion for both into ug/L as N

data_tdkn<-resultconvert(dat = data_tdkn, variable = "Kjeldahl nitrogen", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)
data_tdkn<-detectconvert(dat = data_tdkn, variable = "Kjeldahl nitrogen", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)

data_tdkn<-resultconvert(dat = data_tdkn, variable = "Kjeldahl nitrogen", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)
data_tdkn<-detectconvert(dat = data_tdkn, variable = "Kjeldahl nitrogen", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)

#check to see everything converted 
temp<-data_tdkn %>% filter(CharacteristicName == "Kjeldahl nitrogen")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_tdkn$CharacteristicName<-"Nitrogen, dissolved Kjeldahl"

```


```{r}
#Total Nitrogen

unique(data$CharacteristicName)

TN_params<-c("Nitrogen", "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)") #only values for TN
data_TN<-data[data$CharacteristicName%in%TN_params,] #want to match all records to the vector we created above which have TP
unique(data_TN$ResultSampleFractionText)
temp<-data_TN %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Total")
data_TN<-data_TN[which(data_TN$ResultSampleFractionText %in% fractionkeep | is.na(data_TN$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_TN$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_TN %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #56 NA's - can change
data_TN<-data_TN %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_TN$ResultMeasure.MeasureUnitCode) #checked units
unique(data_TN$DetectionQuantitationLimitMeasure.MeasureUnitCode)

#WORKING WITH "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)" - convert mg/l NO3 to ug/l N = *0.2259*1000

temp<-data_TN %>% filter(CharacteristicName == "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)

data_TN<-resultconvert(dat = data_TN, variable = "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)", unit = "mg/l NO3", out_unit = "ug/l", conv_factor = 0.2259*1000)

data_TN<-resultconvert(dat = data_TN, variable = "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)

data_TN<-detectconvert(dat = data_TN, variable = "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)


temp<-data_TN %>% filter(CharacteristicName == "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#WORKING WITH "Nitrogen" 

temp<-data_TN %>% filter(CharacteristicName == "Nitrogen")
unique(temp$ResultMeasure.MeasureUnitCode) #convert mg/l to ug/l
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)  #all NA's - don't need to convert

  
data_TN<-resultconvert(dat = data_TN, variable = "Nitrogen", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)

temp<-data_TN %>% filter(CharacteristicName == "Nitrogen")
unique(temp$ResultMeasure.MeasureUnitCode) 

data_TN$CharacteristicName<-"Nitrogen, total"

#Completed
```


```{r}
#TDN params - total dissolved nitrogen

unique(data$CharacteristicName)

TDN_params<-c("Nitrogen", "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)") #only values for TN
data_TDN<-data[data$CharacteristicName%in%TDN_params,] #want to match all records to the vector we created above which have TP
unique(data_TDN$ResultSampleFractionText)
temp<-data_TDN %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved")
data_TDN<-data_TDN[which(data_TDN$ResultSampleFractionText %in% fractionkeep | is.na(data_TDN$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_TDN$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_TDN %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #66 NA's - can change
data_TDN<-data_TDN %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_TDN$ResultMeasure.MeasureUnitCode) #checked units
unique(data_TDN$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#WORKING WITH "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)" - convert mg/l NO3 to ug/l N = *0.2259*1000

temp<-data_TDN %>% filter(CharacteristicName == "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)")
unique(temp$ResultMeasure.MeasureUnitCode) #convert mg/l to ug/l
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)  #all NA's - don't need to convert
#need to convert mg/l to ug/l

data_TDN<-resultconvert(dat = data_TDN, variable = "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)

data_TDN<-detectconvert(dat = data_TDN, variable = "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)


temp<-data_TDN %>% filter(CharacteristicName == "Nitrogen, mixed forms (NH3), (NH4), organic, (NO2) and (NO3)")
unique(temp$ResultMeasure.MeasureUnitCode)
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#WORKING WITH "Nitrogen" 

temp<-data_TDN %>% filter(CharacteristicName == "Nitrogen")
unique(temp$ResultMeasure.MeasureUnitCode) #convert mg/l to ug/l
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)  #all NA's - don't need to convert

  
data_TDN<-resultconvert(dat = data_TDN, variable = "Nitrogen", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)

temp<-data_TDN %>% filter(CharacteristicName == "Nitrogen")
unique(temp$ResultMeasure.MeasureUnitCode) 

data_TDN$CharacteristicName<-"Nitrogen, total dissolved"

#Completed


```


```{r}
#ammonium and ammonia


unique(data$CharacteristicName)

NH4_params<-c("Ammonia", "Ammonia and ammonium","Ammonium", "Ammonia-nitrogen") #only values for NH4
data_ammonia<-data[data$CharacteristicName%in%NH4_params,] #want to match all records to the vector we created above which have TP
unique(data_ammonia$ResultSampleFractionText)
temp<-data_ammonia %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved")
data_ammonia<-data_ammonia[which(data_ammonia$ResultSampleFractionText %in% fractionkeep | is.na(data_ammonia$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_ammonia$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_ammonia %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #495 NA's - can change
data_ammonia<-data_ammonia %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_ammonia$ResultMeasure.MeasureUnitCode) #checked units
unique(data_ammonia$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#working with "Ammonia"
temp<-data_ammonia %>% filter(CharacteristicName == "Ammonia")
unique(temp$ResultMeasure.MeasureUnitCode) #convert mg/l to ug/l
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)  #all NA's - don't need to convert

  
data_ammonia<-resultconvert(dat = data_ammonia, variable = "Ammonia", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)

temp<-data_ammonia %>% filter(CharacteristicName == "Ammonia")
unique(temp$ResultMeasure.MeasureUnitCode) 


#working with "Ammonia and ammonium" 

temp<-data_ammonia %>% filter(CharacteristicName == "Ammonia and ammonium")
unique(temp$ResultMeasure.MeasureUnitCode) 
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)  


data_ammonia<-resultconvert(dat = data_ammonia, variable = "Ammonia and ammonium", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)
data_ammonia<-resultconvert(dat = data_ammonia, variable = "Ammonia and ammonium", unit = "mg/l NH4", out_unit = "ug/l", conv_factor = 0.7765*1000)


data_ammonia<-detectconvert(dat = data_ammonia, variable = "Ammonia and ammonium", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)
data_ammonia<-detectconvert(dat = data_ammonia, variable = "Ammonia and ammonium", unit = "mg/l NH4", out_unit = "ug/l", conv_factor = 0.7765*1000)



#working with "Ammonium" # need to finish

temp<-data_ammonia %>% filter(CharacteristicName == "Ammonium")
unique(temp$ResultMeasure.MeasureUnitCode) #convert ueq/L to ug/l
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)  #convert ueq/L to ug/l


data_ammonia<-resultconvert(dat = data_ammonia, variable = "Ammonia and ammonium", unit = "mg/l NH4", out_unit = "ug/l", conv_factor = 0.7765*1000)

data_ammonia<-detectconvert(dat = data_ammonia, variable = "Ammonia and ammonium", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)


#working with "Ammonia-nitrogen"

temp<-data_ammonia %>% filter(CharacteristicName == "Ammonia-nitrogen")
unique(temp$ResultMeasure.MeasureUnitCode) #convert mg/l to ug/l
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)  #convert mg/l to ug/l
#no values - don't need conversion

data_ammonia$CharacteristicName<-"Nitrogen, NH4"

#Completed


```


```{r}

#NO3 params + dissolved inorganic nitrogen

unique(data$CharacteristicName)

NO3_params<-c("Inorganic nitrogen (nitrate and nitrite)", "Nitrate") #only values for DIN
data_NO3<-data[data$CharacteristicName%in%NO3_params,] #want to match all records to the vector we created above which have DIN
unique(data_NO3$ResultSampleFractionText)
temp<-data_NO3 %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved")
data_NO3<-data_NO3[which(data_NO3$ResultSampleFractionText %in% fractionkeep | is.na(data_NO3$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_NO3$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_NO3 %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #1063 NA's - can change
data_NO3<-data_NO3 %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_NO3$ResultMeasure.MeasureUnitCode)
unique(data_NO3$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#working with "Inorganic nitrogen (nitrate and nitrite)"

temp<-data_NO3 %>% filter(CharacteristicName == "Inorganic nitrogen (nitrate and nitrite)")
unique(temp$ResultMeasure.MeasureUnitCode) #convert mg/l and mg/l N to ug/l 
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode) #convert mg/l and mg/l N to ug/l 

  
data_NO3<-resultconvert(dat = data_NO3, variable = "Inorganic nitrogen (nitrate and nitrite)", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)
data_NO3<-detectconvert(dat = data_NO3, variable = "Inorganic nitrogen (nitrate and nitrite)", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)

data_NO3<-resultconvert(dat = data_NO3, variable = "Inorganic nitrogen (nitrate and nitrite)", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)
data_NO3<-detectconvert(dat = data_NO3, variable = "Inorganic nitrogen (nitrate and nitrite)", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)


temp<-data_NO3 %>% filter(CharacteristicName == "Inorganic nitrogen (nitrate and nitrite)")
unique(temp$ResultMeasure.MeasureUnitCode) 
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)


#working with "Nitrate"

temp<-data_NO3 %>% filter(CharacteristicName == "Nitrate")
unique(temp$ResultMeasure.MeasureUnitCode) #need to convert mg/l, mg/l N, mg/lasNO3, and ueq/l to ug/l
unique(temp$DetectionQuantitationLimitMeasure.MeasureUnitCode)  #need to convert mg/l N, mg/lasNO3, and ueq/l to ug/l


data_NO3<-resultconvert(dat = data_NO3, variable = "Nitrate", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)
data_NO3<-resultconvert(dat = data_NO3, variable = "Nitrate", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)
data_NO3<-detectconvert(dat = data_NO3, variable = "Nitrate", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)

data_NO3<-resultconvert(dat = data_NO3, variable = "Nitrate", unit = "mg/l asNO3", out_unit = "ug/l", conv_factor = 0.2259*1000)
data_NO3<-detectconvert(dat = data_NO3, variable = "Nitrate", unit = "mg/l asNO3", out_unit = "ug/l", conv_factor = 0.2259*1000)


###Need to finish

temp<-filter(data_NO3, ResultMeasure.MeasureUnitCode == "ueq/L")
temp<-filter(data_NO3, DetectionQuantitationLimitMeasure.MeasureUnitCode == "ueq/L")

#drop these

data_NO3<-filter(data_NO3, ResultMeasure.MeasureUnitCode != "ueq/L")

data_NO3$CharacteristicName<-"Nitrogen, nitrite (NO2) + nitrate (NO3)"

#completed

```

```{r}
#NO2 params 


unique(data$CharacteristicName)

NO2_params<-c("Nitrite") #only values for DIN
data_NO2<-data[data$CharacteristicName%in%NO2_params,] #want to match all records to the vector we created above which have DIN
unique(data_NO2$ResultSampleFractionText)
temp<-data_NO2 %>% filter(is.na(ResultSampleFractionText)) #to check NA's or anything unclear - we check and decide what to keep
fractionkeep<-c("Dissolved")
data_NO2<-data_NO2[which(data_NO2$ResultSampleFractionText %in% fractionkeep | is.na(data_NO2$ResultSampleFractionText)),] #keeping if it's ones we chose or NA
unique(data_NO2$ResultSampleFractionText)

###create dataset to check for non numeric objects in dataset
temp<-data_NO2 %>% filter(is.na(as.numeric(ResultMeasureValue)))  #checks for non numeric values in result measure value - if NA's not losing any other data - can change #158 NA's - can change
data_NO2<-data_NO2 %>% mutate(ResultMeasureValue = as.numeric(ResultMeasureValue)) # non numeric values become NA

unique(data_NO2$ResultMeasure.MeasureUnitCode)
unique(data_NO2$DetectionQuantitationLimitMeasure.MeasureUnitCode)


data_NO2<-resultconvert(dat = data_NO2, variable = "Nitrite", unit = "mg/l", out_unit = "ug/l", conv_factor = 1000)

data_NO2<-resultconvert(dat = data_NO2, variable = "Nitrite", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)
data_NO2<-detectconvert(dat = data_NO2, variable = "Nitrite", unit = "mg/l as N", out_unit = "ug/l", conv_factor = 1000)

data_NO2<-resultconvert(dat = data_NO2, variable = "Nitrite", unit = "mg/l asNO2", out_unit = "ug/l", conv_factor = 0.3044*1000)
data_NO2<-detectconvert(dat = data_NO2, variable = "Nitrite", unit = "mg/l asNO2", out_unit = "ug/l", conv_factor = 0.3044*1000)


data_NO2$CharacteristicName<-"Nitrogen, nitrite (NO2)"

#completed

```
```


```{r}

alldata_VA<- rbind(data_secchi, data_calcium, data_aluminum, data_mercury, data_magnesium, data_selenium, data_arsenic, data_atrazine, data_ecoli, data_silica, data_chloride, data_oxygen, data_chla, data_pH, data_temp, data_sulfate, data_turb, data_cond, data_carbon, data_alk, data_tss, data_DIP, data_TP, data_TKN, data_TN, data_DIN)     


write.csv(alldata_VA, file = "alldata_VA.csv", row.names = FALSE)

```

